Reconstructing Propositional Proofs in Type Theory

### 2017-10-23 paper/sections/limitations-type-theory.tex
[ paper ] lots of changes.

% ===================================================================

\section{Proof-Reconstruction}
\label{sec:proof-reconstruction}

@@

\subsection{Workflow}
\label{ssec:workflow}

\input{sections/diagram.tex}

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.
%  (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}. The process begin with
a \TPTP file that encodes a problem in \CPL. We use this file as the
input of the \Metis prover. If the problem is a theorem, \Metis
reply a derivation of the proof in \TSTP format.

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
In the \Athena analysis, some unnecessary steps that introduce redundancies
and some unused input is removed from the proof-tree.
As result, we get from \Athena an \Agda file of the proof
with names of functions and theorems from
the \Agda libraries that accompany this article:
\texttt{agda-prop} and \texttt{agda-metis}.

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the proposed formalizations for the
propositional logic and the \Metis inferences.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user looking at the error in the \TSTP derivation
by \Metis, in the translation by \Athena, in the \Agda formalizations
mentioned above or in the type-checker~\Agda.

In the remainder part of this section, a formal description using
type theory is presented to build definitions and theorems of
functions necessary to emulate \Metis' inference rules.

@@

% section section_name (end)
\subsection{Structural recursion}
\label{ssec:structural-recursion}

@@

\begin{example}
\label{ex:structural-recursion}

@@

\end{example}

@@

% % ................................................

\subsection{Reconstructing \Metis' inference rules}
\label{ssec:emulating-inferences}

In this section, we reconstruct each \Metis inference rule.
We present a function and a theorem for each rule.

\begin{notation}
The function name written in bold font is a \Metis inference rule.
The same function name written in roman letters refers to the proposed
version of the rule.
Each rule proposed has a respective theorem also written in roman letters
that proves its soundness.
\end{notation}

The following pattern is used to present the reconstruction.

For example, we know the \verb!conjunct! inference rule is a
\Metis inference rule since it appears in \TSTP derivations like
the following excerpt.

\begin{verbatim}
fof(norm1, (p | r), inf(conjunct, [norma1])).
\end{verbatim}

For this inference rule, we the \rm{conjunct} function in~(\ref{def:conjunct}) and we proved Theorem~\ref{thm:conjunct}. The pattern is the following for
a binary rule and we extend it for $n$-ary rules in a natural way.

\begin{itemize}
\item \conjunct is a binary \Metis inference rule.
\item $\rm{conjunct}$ is the type theory function for the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &\rm{conjunct} : \Prop → \Prop → \Prop → \Prop\\
    &\rm{conjunct}~φ₁~φ₂~ψ\ =
      \begin{cases}
      ψ, &\text{if }\rm{conjunct}\text{ built }ψ\text{ by applying valid transformations}\\
         &\text{or other rules to }φ₁\text{ and }φ₂\\
      φ, &\text{otherwise.}
      \end{cases}
  \end{split}
\end{align*}

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}
\item The theorem that proves the soundness of the
$\rm{conjunct}$ function would be
\begin{equation*}\text{If }Γ ⊢ φ\text{ then }Γ ⊢ \rm{conjunct}\ φ.\end{equation*}
\end{itemize}

\begin{table}[!ht]
\label{tab:agda-metis-table}
\caption{Presentation of \Metis inference rules in type theory.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &Theorem~\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &Theorem~\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &Theorem~\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &Theorem~\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &Theorem~\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &Theorem~\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &Theorem~\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

% -----------------------------------------------------------------------------

\subfile{sections/proof-reconstruction/strip.tex}
\subfile{sections/proof-reconstruction/negate.tex}
\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/resolve.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}

%------------------------------------------------------------------------------

Now, we have finished the formalization of every inference rule in
a \Metis proof, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena jointly with
the \Agda formalizations of the rules mentioned above with more than eighty
representative theorems in propositional logic. An interested reader can test
the problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

>>>>

\section{Limitations of Type Theory}
\label{sec:limitations-type-theory}

@@

@@

% section section_name (end)
\subsection{Terminating functions}
\label{ssec:structural-recursion}

@@

\begin{myexample}
\label{ex:structural-recursion}

@@

\end{myexample}

@@

\subsection{Intuitionistic logic}
\label{ssec:intuitionistic-logc}

...

### 2017-10-23 paper/sections/metis-language-and-proofs.tex
[ paper ] lots of changes.

The \TPTP language --which includes the first-order
form (denoted by \name{fof}) and clause normal form (denoted by
\name{cnf}) formats~\cite{sutcliffe2009,Sicard-Ramirez2016} -- is the
input language to encode problems used by \Metis.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form.

@@

The \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, or a \name{hypothesis}.
The formulas include the binary connectives
(\verb!&!, \verb!|!, \verb!=>!, \verb!<=>!, \verb!~!)
and the constants \verb!$true!, and \verb!$false!.

For instance, let us express the problem
$p\, \vdash \neg (p \wedge \neg p) ∨ (q \wedge \neg q)$
in \TPTP syntax. We begin by declaring the $p$ axiom using the \verb!axiom!
keyword. Later, we include the expected conclusion using the \verb!conjecture!
keyword.

@@

@@

The \name{source} field is an inference record with the following
pattern.

@@

@@

In the script above, \strip is for example the name of the inference.
It has no arguments and derives from one parent node named goal. The
result of this inference when it applies to the \verb!goal! formula is
\verb!p!.

@@

% \TSTP derivations replied by \Metis use other rules. They are
% \canonicalize, \clausify, \conjunct, \simplify, \negate, and \strip.

@@

@@

\begin{figure}
\begin{verbatim}
fof(premise, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p, inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p, inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ∼ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [premise])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [], [normalize_0_0, normalize_0_1]))
cnf(refute_0_0, plain, $false,
  inference(canonicalize, [], [normalize_0_2])).
\end{verbatim}
\caption{\texttt{Metis}' \texttt{TSTP} derivation for the
problem $p\vdash p$.}
\label{fig:metis-proof-tstp}
\end{figure}

The output of a proof delivered by \Metis encodes a natural
deduction proof (see a complete proof in Fig.~\ref{fig:metis-proof-tstp}.
The deduction system uses six inference rules \cite{hurd2003first}.
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (e.g.,
\emph{falsium} in the last conclusion in the \TSTP derivation
Fig.~\ref{fig:metis-proof-tstp}).

@@

\begin{figure}[!ht]
\centering
  \begin{bprooftree}\tt
    \AxiomC{}
    \RightLabel{negate}
    \UnaryInfC{$\neg p$}
    \RightLabel{strip}
    \UnaryInfC{$\neg p$}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$p$}
    \RightLabel{canonicalize}
    \UnaryInfC{$p$}
    \RightLabel{simplify}
    \BinaryInfC{$⊥$}
    \RightLabel{canonicalize}
    \UnaryInfC{$⊥$}
  \end{bprooftree}
  \caption{Proof tree for the \Metis derivation in
  Fig.~\ref{fig:metis-proof-tstp}}.
  \label{fig:metis-example}
\end{figure}

>>>>

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal form (denoted by \name{cnf}) formats~\cite{sutcliffe2009,Sicard-Ramirez2016}.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form:

@@

Where the \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} or an \name{hypothesis}.
The formulas include the constants \verb!$true!, negation unary operator (\verb!~!), and the binary connectives (\verb!&!, \verb!|!, \verb!=>!, \verb!<=>!),

\begin{myexample}
For instance, let us express the problem
$p\, \vdash \neg (p \wedge \neg p) ∨ (q \wedge \neg q)$
in \TPTP syntax. We begin by declaring the $p$ axiom using the \verb!axiom!
keyword. Later, we include the expected conclusion using the
\verb!conjecture! keyword.

@@

\end{myexample}

@@

Where the \name{source} field is an inference record with the following
pattern:

@@

\begin{myexample}

@@

In the script above, \strip is for example the name of the inference.
It has no arguments and derives from one parent node named  \verb!goal!. The
result of this inference when it applies to the \verb!goal! formula is
\verb!p!.

\end{myexample}

We adopt a customized \TSTP syntax to keep as short as possible the \Metis derivations for increasing the readability in this paper.

\begin{myexample}

Let us consider the following \TSTP derivation using
the customized \TSTP syntax. Note that we remove for example
empty fields in the inference information, the \verb!role! field if it is
\verb!plain! and shorten names as long as possible.

\begin{verbatim}
fof(premise, axiom, p).
fof(goal, conjecture, p).
fof(sg0, p, inf(strip, [goal])).
fof(neg0, ~ p, inf(negate, [sg0])).
fof(norm0, ∼ p, inf(canonicalize, [neg0])).
fof(norm1, p, inf(canonicalize, [premise])).
fof(norm2, $false, inf(simplify, [norm0, norm1]))
cnf(refute0, $false, inf(canonicalize, [norm2])).
\end{verbatim}

\end{myexample}

@@

% \TSTP derivations replied by \Metis use the following rules:
% \canonicalize, \clausify, \conjunct, \simplify, \negate, and \strip.

@@

\begin{table}[!ht]
\label{tab:agda-metis-table}
\caption{Presentation of \Metis inference rules in type theory.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &Theorem~\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &Theorem~\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &Theorem~\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &Theorem~\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &Theorem~\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &Theorem~\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &Theorem~\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

@@

A proof delivered by \Metis encodes a natural
deduction proof (see a complete proof in Fig.~\ref{fig:metis-proof-tstp}).
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (\eg,
\emph{falsium} in the root of the \TSTP derivation
in Fig.~\ref{fig:metis-proof-tstp}).

@@

\begin{figure}[!ht]
\centering
  \begin{bprooftree}\tt
    \AxiomC{}
    \RightLabel{negate}
    \UnaryInfC{$\neg p$}
    \RightLabel{strip}
    \UnaryInfC{$\neg p$}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$p$}
    \RightLabel{canonicalize}
    \UnaryInfC{$p$}
    \RightLabel{simplify}
    \BinaryInfC{$⊥$}
    \RightLabel{canonicalize}
    \UnaryInfC{$⊥$}
  \end{bprooftree}
  \caption{A \Metis proof tree.}
  \label{fig:metis-example}
\end{figure}

### 2017-10-23 paper/sections/proof-reconstruction.tex
[ paper ] Eq removed.
[ paper ] last changes.
[ paper ] lots of changes.

MISSING TEXT

@@

\input{sections/diagram.tex}

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.
%  (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}. The process begin with
a \TPTP file that encodes a problem in \CPL. We use this file as the
input of the \Metis prover. If the problem is a theorem, \Metis
reply a derivation of the proof in \TSTP format.

@@

\begin{notation}
We write the type judgments as $m : \abbre{M}$ to denote
that $m$ has $\abbre{M}$ type.
Types are written using small caps.
We use \Prop for the type of propositions and the dependent type, $Γ ⊢ φ$ for
theorems where $Γ : \List~\Prop$ and $φ : \Prop$.
$\abbre{List}$ type is the usual inductive type for lists.
For syntactical equivalence, we write the equality as $φ ≡ ψ$
for $φ,~ψ : \Prop$. For this equality, we assume the
reflexivity, symmetry and transitivity properties.
\end{notation}
% ...................................................................

% section section_name (end)
\subsection{Structural recursion}
\label{ssec:structural-recursion}

To reconstruct \Metis' inference rules in type theory, we observed that
transformations performed by some rules are implemented by \emph{general
recursive} functions.
Algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate. For that
reason, we follow the technique described in~\cite{Bertot2004} to avoid
termination problems by modifying the recursive functions to be
\emph{structural recursive}.
% (see \cite{Coquand1992,Abel2002,Bove2005} for more details).

This paper adopts the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
structurally smaller arguments.

General recursive functions can be translated into structurally recursive
functions by using for instance, the \emph{Bounded Recursion} technique
presented in~\cite{Bertot2004}. For a yet another methods, we refer
the reader to \cite{Coquand1992,Abel2002,Bove2005}.
The bounded technique defines a new structural recursive function based on
the general recursive function by adding an argument to the function.
The new argument is the \emph{bound}, a natural number given by the
function complexity. In other words, the added argument will represent
the number of times the function needs to call itself to get the
expected outcome~\cite{Bertot2004}.

\begin{notation}
The type of natural numbers is called \Nat, and it is defined as
usual, where \zero and \suc are its data constructors. We use names
and symbols of the arithmetic operations as usual.
We include syntax sugar for $\zero$, $\suc$, $\suc(\zero)$,
$\cdots$, with the decimal representation: $0$, $1$, $2$, $\cdots$
as well.
\end{notation}

Therefore, one approach to guarantee a structural recursive function based on
a general recursive function $\rm{f}: \Prop \to \Prop$ is to formulate a new function $\rm{f}^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls are done on smaller arguments.

\begin{example}
\label{ex:structural-recursion}

Let us consider the following example to exhibit the bounded technique
mentioned above for the \fuh function. this function is used for reconstructing
the \Metis \strip inference rule in Section~\ref{sssec:strip-a-goal}.

\begin{equation}
\label{eq:uh-definition}
\begin{aligned}
&\hspace{.495mm}\fuh_{0} : \Prop → \Prop\\
&\begin{array}{lll}
  \fuh_{0} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{0}~((φ₁ ∧ φ₂) ⇒ φ₃)\\
  \fuh_{0} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fuh_{0}~(φ₁ ⇒ φ₂) ∧ \fuh_{0}~(φ₁ ⇒ φ₃)\\
  \fuh_{0} &φ                &= φ
\end{array}
\end{aligned}
\end{equation}

In (\ref{eq:uh-definition}), the first two equation show
call on arguments no structurally smaller.
For instance, on the first call, $φ ≡ (φ₁ ∧ φ₂) ⇒ φ₃$, the recursion
calls on the  $φ₁ ⇒ (φ₂ ⇒ φ₃)$ argument.
But this argument is not a subformula for φ.
Therefore, one way to translate $\fuh_{0}$ in (\ref{eq:uh-definition}) to type
theory is to define the a new function using the bounded technique.
\begin{equation}
\label{eq:uh-structured}
\begin{aligned}
&\hspace{.495mm} \fuh_{1} : \Prop → \Nat \to \Prop\\
&\begin{array}{llll}
\fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
\fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
\fuh_{1} &φ &\xspace\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

As we mentioned above, the natural number in the second argument of
the function stands for a complexity measure of it.

Recall the \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number for a proposition $φ$ like the following $\frank$ function.
The symbols $c_{¬}$ and $c_{\square}$ are non-negative integer
numbers called the complexity of the respective connective.

\begin{equation}
% \label{eq:rank-definition}
\begin{aligned}
&\hspace{.495mm}\frank : \Prop → \Nat\\
&\begin{array}{llll}
\frank~(¬~φ)           &= \frank~φ₁~+~c_{¬}\\
\frank~(φ₁~\square~φ₂) &= \fmax~(\frank~φ₁)~(\frank~φ₂) + c_{\square}\\
\frank~φ               &= 0
\end{array}
\end{aligned}
\end{equation}

We define the $\fuh_{cm}: \Prop → \Nat$ function as
the complexity measure for the $\fuh_0$ function.

\begin{equation*}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0
    \end{array}
  \end{aligned}
\end{equation*}

Following the technique mentioned above, we get the structural recursive definition of $\fuh_{0}$ function presented in ~(\ref{eq:uh}).

\begin{align}
  \label{eq:uh}
  \begin{split}
  &\fuh : \Prop → \Prop\\
  &\fuh~φ~ = \fuh_{1}~φ~(\fuh_{cm}~φ).
  \end{split}
\end{align}

\end{example}

For a more complete treatment of general recursion, and how to
define well-founded and well-structural recursive functions, we refer
the reader to \citeauthor{Bove2005}~\cite{Bove2005}.

@@

In this section, we reconstruct each \Metis inference rule.
We present a function and a theorem for each rule.

@@

The following pattern is used to present the reconstruction.

For example, we know the \verb!conjunct! inference rule is a
\Metis inference rule since it appears in \TSTP derivations like
the following excerpt.

\begin{verbatim}
fof(norm1, (p | r), inf(conjunct, [norma1])).
\end{verbatim}

For this inference rule, we the \rm{conjunct} function in~(\ref{def:conjunct}) and we proved Theorem~\ref{thm:thm-conjunct}. The pattern is the following for
a binary rule and we extend it for $n$-ary rules in a natural way.

\begin{itemize}
\item \conjunct is a binary \Metis inference rule.
\item $\rm{conjunct}$ is the type theory function for the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &\rm{conjunct} : \Prop → \Prop → \Prop\\
    &\rm{conjunct}\ φ₁\ φ₂\ ψ\ =
      \begin{cases}
      ψ, &\text{if }\rm{conjunct}\text{ built }ψ\text{ by applying valid transformations}\\
         &\text{or other rules to }φ₁\text{ and }φ₂\\
      φ, &\text{otherwise.}
      \end{cases}
  \end{split}
\end{align*}

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}
\item The theorem that proves the soundness of the
$\rm{conjunct}$ function would be
\begin{equation*}\text{If }Γ ⊢ φ\text{ then }Γ ⊢ \rm{conjunct}\ φ.\end{equation*}
\end{itemize}

\begin{table}[!ht]
\label{tab:agda-metis-table}
\caption{Presentation of \Metis inference rules in type theory.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &Theorem~\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &Theorem~\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &Theorem~\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &Theorem~\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &Theorem~\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &Theorem~\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &Theorem~\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

MISSING TEXT

>>>>

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.

@@

\input{sections/diagram.tex}
%  (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}. The process begin with
a \TPTP file that encodes a problem in \CPL. We use this file as the
input of the \Metis prover. If the problem is a theorem, \Metis
reply a derivation of the proof in \TSTP format.

@@

@@

In this section, we reconstruct each \Metis inference rule with a
function and its respective theorem. We present a pattern of the
formal description for these rules.

@@

\begin{myexample}
\Metis has a inference rule named conjunct.
Since that rule appears in \TSTP derivations like in the following excerpt.
We refer to this rule writing \conjunct.

\begin{verbatim}
fof(p3, (r | ~ p), inf(conjunct, [p1])).
\end{verbatim}

The pattern of the formalization for this rule is the following.
For \conjunct rule, we define in type theory a function using the same name,
conjunct, but written using roman font to distinguish from the original.
The \fconjunct function has the following pattern:

 \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fconjunct : \Prop → \Prop → \Prop\\
  &\begin{array}{ll}
  \fconjunct~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if }\fconjunct \text{ built }ψ\text{ by applying valid}\\
         &\text{transformations or other rules to }φ\\
      φ, &\text{otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}

\begin{itemize}
  \item[$\bullet$] $φ$ is the \emph{source} formula from a previous deducted derivation.
  \item[$\bullet$] $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}

To justify all transformations done by \fconjunct, we prove its soundness
with a theorem. This theorem would be

\begin{equation*}
\text{If }Γ ⊢ φ\text{ then }Γ ⊢ \fconjunct~φ~ψ.
\end{equation*}

\end{myexample}

The remainder of this section will be devoted to present a formal description
in type theory of each rule presented in Table~\ref{tab:agda-metis-table} using
the pattern present in the example above.

### 2017-10-23 paper/sections/proof-reconstruction/negate.tex
[ paper ] lots of changes.

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal.
The \negate rule assumes the negation of such subgoals. Then, this rules
is always derived from a strip inference application as we can see in 
the following \TSTP excerpt.

@@

To emulate this rule, we only need to introduce the negation
of the subgoal by using $assume$ rule from the inferences in
Fig.~\ref{fig:CPL-inference-rules}.

>>>>

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal.
The \negate rule assumes the negation of such subgoals. Then, this rules
is always derived from a \strip inference application as we can see in
the following \TSTP excerpt.

\begin{myexample}
\negate is a rule that only puts the negation symbol
in front of the formula to assume it.

@@

\end{myexample}

To emulate this rule, we only need to introduce the negation
of the subgoal by using $assume$ rule.
